%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The 2nd Trial by Andrea Tino is licensed under CC BY-NC-SA 4.0. To view a   %
% copy of this license, visit                                                 %
% http://creativecommons.org/licenses/by-nc-sa/4.0/                           %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter[Questions]{Questions}
\label{ch:qs}

In this year's exams we can see a nice variaty of topics in the questions
ranging from the main theme: Magnetism to Probability and Linear Algebra.

\section{Extrema finding}
\label{sec:extremafind}
\marginnote{Appears as problem 1 in L 06.}
Let $p(x)$ be a polynomial, the following function is provided:
\begin{equation*}
    f(x) = \frac{p(x)}{x^2 + d}
\end{equation*}
Where $d \in \mathbb{R}$. $f$'s plot intersects the x-axis at $x_1 = 0$ and
$x_2 = \frac{12}{5}$, and has these lines as asymptotes: $x=3$, $x=-3$ and
$y=5$. We need to calculate the points of relative maximum and minimum.

\paragraph{}
The first thing important noticing is that $f$ is the ratio between two
polynomials. How much do we know about these two polynomials? The
denominator is a second-degree polynomial, while we do not have much
information about the numerator. We need to know a few things more about
$p$. From what the problem tells us, we know that
$\lim_{x \to \infty} f(x) \neq \infty$. That is a very important piece
of data because we know that the ratio between two polynomials tends
to a finite value (as the variable tends to infinity) only when they
have the same
degree\sidenote{The full theorem actually states that:
\begin{equation*}
    \lim_{x \to \infty} \frac{N(x)}{D(x)} = l \neq \infty
        \iff n = m
\end{equation*}
Where $N(x)$ and $D(x)$ are two polynomials with
degrees $n$ and $m$ respectively. And we also have that:
\begin{equation}
    l = \frac{a_N}{a_D}
\end{equation}
Where $a_N$ id $N$'s highest power's coefficient, and
$a_D$ id $D$'s highest power's coefficient.}.
Thanks to it, we can rewrite $f$ as:
\begin{equation}\label{eq:fracinfty}
    f(x) = \frac{ax^2 + bx + c}{x^2 + d}
\end{equation}
Because of equation \ref{eq:fracinfty}, we have that:
\begin{equation*}
    \lim_{x \to \infty} f(x) = \lim_{x \to \infty} \frac{ax^2 + bx + c}{x^2 + d}
        = \frac{a}{1} = 5
        \implies a = 5
\end{equation*}
As we unveil the values of coefficients, we know more and more the final form
of $f$, and this is important because we will need to compute its derivative and
then study its sign: the less parameters in the derivative, the easier our job.
Now $f$ is:
\begin{equation*}
    f(x) = \frac{5x^2 + bx + c}{x^2 + d}
\end{equation*}
We still have 3 parameters.
Let's use the info about intersections with the x-axis to find the values of some
of those:
\begin{equation*}
    \begin{split}
        &\begin{cases}
            f(0) = 0\\
            f\left( \frac{12}{5} \right) = 0
        \end{cases}
        \implies
        \begin{cases}
            \frac{c}{d} = 0\\
            \frac{5\frac{12^2}{5^2} + b\frac{12}{5} + c}{\frac{12^2}{5^2} + d} = 0
        \end{cases}
        \implies
        \begin{cases}
            c = 0\\
            \frac{\frac{12^2}{5} + b\frac{12}{5}}{\frac{12^2}{5^2} + d} = 0
        \end{cases}
        \implies\\
        &\frac{12^2 + 12b}{5} = 0 \implies
        12 + b = 0 \implies
        b = -12
    \end{split}
\end{equation*}
Which means now $f$ can be written as:
\begin{equation*}
    f(x) = \frac{5x^2 - 12x}{x^2 + d}
\end{equation*}
We only have $d$ left to find out, and for that we can use the information about
the two vertical asymptotes. Remember that $f$ is a ratio, when the denominator
tends to 0, the whole ratio tends to infinity; it means that if we find the
values of $x$ for which $x^2 + d = 0$, we will find the exact points where
the asymptots are: $x^2 + d = 0 \implies x = \pm \sqrt{-d}$. At this point
it becomes clear that $d < 0$ because otherwise we would have no asymptotes as
we would need to extract the square root of a negative number which is not a
real number.
Also notice that $d \neq 0$, otherwise we would not have two different
asymptotes, but only one. Since we know that the two aymptotes are $x = \pm 3$,
we can conclude that: $\pm 3 = \sqrt{-d} \implies d = -9$, and:
\begin{equation*}
    f(x) = \frac{5x^2 - 12x}{x^2 - 9}
\end{equation*}
Now we can compute $f^\prime$:
\begin{equation*}
    f^\prime(x) = \frac{10x - 12}{x^2 - 9} - \frac{5x^2 - 12x}{(x^2 - 9)^2} =
        \frac{10x^3 - 17x^2 - 78x + 108}{(x^2 - 9)^2}
\end{equation*}
We first need to find the zeros\sidenote{Remember that the zeros
of a fraction $\frac{N(x)}{D(X)}$ can be found by finding the zeros of the numerator $N(x)$.}
of $f^\prime$:
\begin{equation}\label{eq:pol3roots}
    f^\prime(x) = 0 \implies 10x^3 - 17x^2 - 78x + 108 = 0
\end{equation}
Here's the bad news: $f^\prime$ is a complete\sidenote{A cubic
polynomial $ax^3 + bx^2 + cx + d$ is \textit{complete} when all coefficients are
present: $a,b,c,d \neq 0$.} cubic polynomial; the methods for finding its exact roots
are quite advanced and also very calculation-intensive. We don't wanna go down that road.
Here's the good news, we notice that all coefficients are integer numbers; this is
good because we could use the
\smallcaps{Rational Root Theorem}\sidenote[][-0.05\textwidth]{The \textit{Rational Root Theorem} states that
if an $n$-order polynomial $P(x)$:
\begin{equation*}
    P(x) = a_x x^n + a_{n-1}x^{n-1} + \dots + a_0
\end{equation*}
has a rational root $x^\ast$:
\begin{equation*}
    P(x^\ast) = 0 \wedge x^\ast = \frac{p}{q} \wedge p,q \in \mathbb{N}
\end{equation*}
Then that root can be found inside $\Omega_P$:
\begin{equation*}
    \Omega_P \equiv \left\{ \frac{p_i}{q_j} : p_i \in \text{m}(a_0) \wedge q_j \in \text{m}(a_n) \right\}
\end{equation*}
Where $\text{m}(\alpha)$ indicates all the prime factors of integer $\alpha$.}.
If you haven't studied such theorem,
the only feasible way to solve \ref{eq:pol3roots} is using numerical methods like
we did in \nameref{par:subp6par1_1}.

\section{A big polynomial}
\label{sec:bigpoly}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Parallelepipeds}
\label{sec:paralpyd}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Spheres}
\label{sec:speheres}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Tossing dices}
\label{sec:dices}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Magnetic fields and amperian loops}
\label{sec:magnamp}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Spaceships}
\label{sec:spaceship}
\marginnote{Appears as problem 1 in L 06.}
Something

\section{Protons in motion}
\label{sec:protons}
\marginnote{Appears as problem 1 in L 06.}
Something
